---
title: Week 12 Worksheet
subtitle: Comparing more than two group means 
output:
  html_document
---


## Part 1 

Simulate 7 batches of 10 normally distributed random numbers with mean 4 and variance 0.0037.  

```{r}
my_data <- data.frame(batch = factor(rep(1:7, 10)), measurement = rnorm(70, 4, sqrt(0.0037)))
summary(my_data)
```

## Part 2

Edit the title below to make a personalized parallel box plot of these seven batches and comment on what you see. 

```{r}
library(ggplot2)

ggplot(data=my_data, aes(x=batch, y=measurement)) + 
  geom_boxplot() + ylim(3.7,4.3) + 
  labs(title = "[Your Name]'s Plot", x = "batch", y = "measurement")
```

## Part 3

Upload your plot image to [this folder](https://drive.google.com/drive/folders/11mZv7IlA9wbYPn0HsVfdcHedMOcXqXQh?usp=sharing).

Take a look at the other plots in this folder before moving on to Part 4. 


## Part 4

* Do any plots appear to have differences in the group variances? 

* Do any plots appear to have differences in the group means? 


If there's still time remaining while others finish, feel free to work on Part 5. 


## Part 5

Add a second factor to your data set that has 5 levels. Use the code below to create plots of the treatment means. 

```{r}
my_data2 <- data.frame(lab = factor(rep(c("A", "B", "C", "D", "E"), 14)), my_data)

ggplot(data = my_data2, aes(x=batch, y=measurement, color=lab, group=lab)) + 
  stat_summary(fun = mean, geom="point") + 
  stat_summary(fun = mean, geom="line") +
  labs(title = "[Your name]'s Interaction Plot")
```

* What do the different line colors represent? 

* Would you recommend using an additive model or an interaction model? 

