A college campus tests the processing times for 31 randomly selected desktop computers. Some computers are Mac desktop computers, the others are Dell PC computers. Each computer runs the same code and the processing time for this code is recorded. Is the average processing time of the Mac desktops different from that of the Dell desktops? 

 A random sample of 35 van drivers from two different nation-wide moving business measures the total number of miles driven during a fixed, nine-week period. Is there a statistically detectable difference in the average miles driven between the two businesses? 

For two types of smudge pots used to protect orchards from frost, the burning time (in minutes) is recorded. Is there a difference in the average burn time for these two types of smudge pots? 

he professional staff at a 600-bed hospital conduct a study of the variability of the weights recorded by beam scales at their facility for 65 randomly selected scales. Two certified test weights of 50 lbs and 150 lbs are weighed on each scale in a randomly selected order and the absolute differences between the certified weight and the beam scale reading are recorded. Are the hospital's beam scales more precise for heavier weights? 

In a large city, two individual stores of a large chain are thought to have equally attractive locations. For 60 randomly selected business days, the total daily sales for each store is recorded. Is there a difference in the expected sales between the stores? 


For which of the statistical tests above are we likely to need to control for multiplicity?

For each of the scenarios above, select a significance level (from a choice of options) and justify decision. 



TO DO: create edited/fixed copies of the quizzes for re-submission, must follow instructions and answer every problem entirely (even if you got the problem correct the first time) fresh slate, tabula rasa. 